---
title: "Workshop2"
author: "Sanjana Rambojun-Ruchaya"
date: "2025-09-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Install and load tidyverse packages
```{r}
## install.packages("tidyverse") # Delete this line once installed
library("tidyverse")
```

# Load the data
```{r}
data(mpg)

```

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

#Labels
```{r}
library(ggplot2)

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(title = "Fuel efficiency generally decreases with engine size")
```

```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  labs(
    title = "Fuel efficiency generally decreases with engine size",
    subtitle = "Two seaters (sports cars) are an exception because of their light weight",
    caption = "Data from fueleconomy.gov"
  )
```
```{r}
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_smooth(se = FALSE) +
  labs(
    x = "Engine displacement (L)",
    y = "Highway fuel economy (mpg)",
    colour = "Car type"
  )
```
#Annotations
```{r}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 2)

#Nudge the text labels up a little to avoid overplotting (nudge_x or y)
#Can use position= position_jitter(width=, height=) to avoid overplotting
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class, nudge_y = 2)
```
#Scales
```{r}
#?scale_x_continuous()  #?scale_y_continuous() #?scale_colour_discrete()
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  scale_x_continuous() +
  scale_y_continuous() +
  scale_colour_discrete()
```
#Axis ticks
```{r}
#Breaks controls ticks' positions and you can use labels as an argument to change the text label associated with ticks. 
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_y_continuous(breaks = seq(15, 40, by = 5))
```
```{r}
#Can use labels set to NULL to suppress the labels altogether.
ggplot(mpg, aes(displ, hwy)) +
  geom_point() +
  scale_x_continuous(labels = NULL) +
  scale_y_continuous(labels = NULL)
```
#Legends and colour schemes
```{r}
base <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class))

base + theme(legend.position = "left")
base + theme(legend.position = "top")
base + theme(legend.position = "bottom")
base + theme(legend.position = "right") # the default
base + theme(legend.position = "none")
```
#Replacing a scale
```{r}
#Log transform using scale functions
ggplot(diamonds, aes(carat, price)) +
  geom_bin2d() + 
  scale_x_log10() + 
  scale_y_log10()
```
```{r}
#Use scale functions to change the colour scheme
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv))

ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv)) +
  scale_colour_brewer(palette = "Set1")
```

```{r}
#Redundant shape mapping
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = drv, shape = drv)) +
  scale_colour_brewer(palette = "Set1")
```

```{r}
#Custom colour scheme
presidential %>%
  mutate(id = 33 + row_number()) %>%
  ggplot(aes(start, id, colour = party)) +
    geom_point() +
    geom_segment(aes(xend = end, yend = id)) +
    scale_colour_manual(values = c(Republican = "red", Democratic = "blue"))
```

```{r}
#Viridis and hexbin colour schemes
install.packages('viridis')
install.packages('hexbin')
library(viridis)
library(hexbin)
```

```{r}
df <- tibble( # note we're just making a fake dataset so we can plot it
  x = rnorm(10000),
  y = rnorm(10000)
)
ggplot(df, aes(x, y)) +
  geom_hex() + # a new geom!
  coord_fixed()

ggplot(df, aes(x, y)) +
  geom_hex() +
  viridis::scale_fill_viridis() +
  coord_fixed()
```

#Themes
```{r}
#Theme 1
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_bw()

#Theme 2
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_light()

#Theme 3
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_classic()

#Theme 4
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme_dark()

#Custom theme
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme (panel.border = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position="bottom",
        legend.title=element_blank(),
        legend.text=element_text(size=8),
        panel.grid.major = element_blank(),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.text.y=element_text(colour="black"),
        axis.text.x=element_text(colour="black"),
        text=element_text(family="Arial"))
#Use sans instead of Arial after checking fonts available in system
```
#Checking Windows fonts
```{r}
windowsFonts()
#To see all the fonts available on your system
```
```{r}
#Using sans font
ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(color = class)) +
  geom_smooth(se = FALSE) +
  theme (panel.border = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.position="bottom",
        legend.title=element_blank(),
        legend.text=element_text(size=8),
        panel.grid.major = element_blank(),
        legend.key = element_blank(),
        legend.background = element_blank(),
        axis.text.y=element_text(colour="black"),
        axis.text.x=element_text(colour="black"),
        text=element_text(family="sans"))
```
#Saving and exporting plots
```{r}
ggplot(mpg, aes(displ, hwy)) + geom_point()

ggsave("my-plot.pdf")
#> Saving 7 x 4.32 in image
```

